Ontology Matching with word2vec

### 2016-04-12 Paper/Draft.tex
Added Short context, method and some newpages

 \begin{document}
 \maketitle
 \newpage

@@

 \section{Introduction}
 It is a Friday afternoon and you realise you want to watch a movie with some friends. So, what do you do? To achieve this goal you must take 4 steps.

@@

@@

\section{Context}
% Give an overview of the section
 In this section I will discuss all concepts that are required to understand the study. Firstly, the Semantic Web, the key technology that depends very strongly on ontology matching, will be discussed. Then the concept of the Ontology itself will be addressed. Lastly, current alignment strategies, their problems and other research and technologies will be discussed.
 \subsection{The Semantic Web}
 The Semantic Web, or Web 3.0, is a concept of a Web that enhances the User Experience in ways that the current Web can not provide. Web 2.0, the current Web, models the internet as \textit{hubs} (or sites) with a specific purpose and barely any interconnectivity. For example, Facebook is a hub for sharing content with people you consider friends, but if you want to share that content with more people you will have to go to another hub, like imgur or Pinterest.
The Semantic Web, on the other hand, sees web pages as sources of \textit{information} that can be combined whenever the user needs it. This allows not only humans to browse the internet, but also Artificial Intelligences, which will use this information to aid their users.

@@

 \subsection{Other related research}
 \subsubsection{Word sense representation}
 Skip-gram models have been widely used to represent words as vectors. This is useful as it is easy to calculate the distance between two vectors, which is a useful property for ontology matching. This vector distance has been shown to be related to the semantic similarity between words, which is very useful for ontology alignment, as parts of one ontology have to be aligned with parts of another ontology depending on their semantic relation.

@@

 Word2vec has been expanded in a number of different ways. For example, word order can be preserved, leading to a similarity measure that is closer to syntax, as syntax defines which words go where in a sentence. This preservation was done by having multiple output layers, one for every word in the context. The research shows that the input does not have to be a bag-of-words representation, which will be useful later on. \cite{ling2015}
 Another interesting development is the multi-sense word2vec, which allows for multiple vector representations per word, depending on the number of different definitions a word has. For example, the word bank represents both the monetary institute and the riverside, which would both have a different representation in this extension. It can differentiate between two different meanings based on the context. This will help with resolving ambiguity, a very important factor in ontology alignment.\cite{multisense}
 \subsection{Used technology}
 \subsubsection{AgreementMaker}
 AgreementMaker is an ontology matching system which obtained the highest F-measure in 6 of the 7 ontology matching tracks of OAEI 2015. \cite{OAEI2015} The system has also been updated the most recent at the time of writing\cite{agreementmaker}.

 The system uses an extensible framework which allows its users to add new modules to the system. This is very useful for research as it allows the researcher to compare different modules by swapping in just those modules in the framework. The module-swapping technique will also be used in this work, but more of that will be discussed in \ref{Evaluation methods}.
 \section{Problem}

 \subsection{Problem}
 We have seen that when information from multiple domains needs to be integrated so that it is possible to reason over them together, the ontologies of those domains need to be aligned or merged. % Waar?
  Those ontologies contain rules about concepts and concept classes in the shape of triplets, with two concepts and a relation. For example, the rule{\tt Professors are humans} contains the concepts \emph{Professor} and \emph{human} and relationship \emph{being}.

@@

 In summary, ontology aligning is required for multi-domain communication but poses a number of problems that make it difficult for concepts, the building blocks of ontologies, to be matched.
 \subsection{Research questions}
 Can distributed representations of concepts be used to find relationships between concepts in ontologies better than existing terminological matchers?

 Can distributed representations use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers?

@@

 \subsection{Hypotheses}
 Since multiple implementations of the algorithm are considered, every hypothesis that follows is considered for every implementation.

 \textit{The distributed representations of concepts matcher can find relationships between concepts in ontologies better than existing terminological matchers.}

 \textit{The distributed representations matcher can use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers.}

 \textit{The distributed representations matcher can perform well in a domain it was not designed for.}

@@

 \textit{A distributed representations of concepts matcher can find relationships between concepts in ontologies better than existing terminological matchers.}

 \textit{A distributed representations matcher can use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers.}

 \textit{A distributed representations matcher can perform well in a domain it was not designed for.}

@@

 However, ontology matching is also used outside of the semantic web. It is already in use for companies that want to merge knowledge from other companies, for example. If the proposed matching algorithm is better than the current algorithms, or at least improves the performance when used in ensemble with the current algorithms, it will be useful in any situation, not just on the Semantic Web.
 \section{Method}
 In this section, I will describe the method that I developed. Firstly, I will describe the general idea. In the second section, I will give a theoretical justification for why this method should work and how it works conceptually. After that I will describe the different ways that the input graphs can be used to train the word2vec model and create the alignment. 
 \subsection{Idea}
 As explained earlier, word2vec is a model that learns two things: implicitly it learns the semantic vector representation of words, and explicitly it learns to predict words from their context or vice versa. I will use both of these pieces of information in our alignment algorithm.
 The semantic representation will be used for finding nodes in the two ontology graphs that are similar, i.e. close in the vector space. This step is relatively straightforward if proper vector representations are found. However, this may be hard due to a number of problems which I have listed below.

@@

 \subsection{Discarded possibilities} % Rename

@@

 \subsection{Implementation}

 \section{Evaluation}
 %https://hal.inria.fr/hal-00917910/document
 The goal of the research is to find an ontology alignment algorithm $A$ and proof that it performs better than current techniques. However, as current techniques may combine many different algorithms and sources of information, maybe such a comparison is not fair, nor is a comparison with individual algorithms as they may be optimised for being combined. So, instead we may want to see if performance is improved when adding $A$ to an ensemble of algorithms $E=(A_1,..,A_n)$.
 \subsection{Evaluation methods} \label{Evaluation methods}
 The evaluation should be done using a performance measure $P(Al,T)$ that takes the alignment an algorithm produces given an ontology and the ground truth. If both the alignment and truth are considered sets of alignments, i.e. $Al=\{a_1,..,a_n\}$ and $T=\{a'_1,..,a'_m\}$, then the True Positives $TP=\{a|a\in Al, a\in T\}$, False Positives $FP=\{a|a\in Al, a\notin T\}$ and False Negatives $FN=\{a|a\in T, a\notin Al\}$ determine the performance of the alignment. The True Negatives do not matter, as there are extremely many, and the goal of ontology matching is to find True Positives. Therefore the Jaccard similarity should be used as a performance measure.

@@

 \section{Algorithm}
 \subsection{Idea}
 The goal of any ontology matcher is to find an alignment between two ontologies that is as similar as possible to the true alignment. Such an alignment consists of node pairs where the two nodes in each pair come from the two different ontologies. Sometimes a label is added to each pair, indicating the relationship the two nodes have. For example, 'part-of', 'similar-to', and such. However, we only consider equality relationships and thus the label can be omitted.

@@

 Word2vec is an unsupervised method that maps words to a vector space that has been shown to contain semantic properties %cite.
 It can be trained on large corpora quickly and the resulting vectors can easily be used to measure the distance between words. All these properties make it an excellent candidate for a concept distance measure to use for ontology matching. As an added benefit, it can also take into account the context of a concept, which may be able to improve its performance over a matcher that only uses the label of a concept.
 \subsection{Theory}

@@

 \section{Results} \label{results}

 \section{Conclusions}
 \subsection{Recap research questions}
 \subsection{Recap hypotheses}
 \subsection{Conclusion per hypothesis}
 \subsection{Overview of results}
 \subsection{Final conclusion}

 \section{Discussion}
 \subsection{Interpretation}
 \subsection{Unexpected results}
 \subsection{Expected results}
 \subsection{Future research}

 \begin{appendices}
  \section{Appendix 1}
  \section{Appendix 2}
 \end{appendices}

 \section*{References}
 \bibliographystyle{plain}
 \bibliography{bibliography}{}
 \end{document}

>>>>

 \begin{document}
 \pagenumbering{gobble}
 \maketitle
 \newpage

 \pagenumbering{roman}
 \section*{Acknowledgements}
 \section*{Abstract}

@@

 \pagenumbering{arabic}
 \section{Introduction}
 It is a Friday afternoon and you realise you want to watch a movie with some friends. So, what do you do? To achieve this goal you must take 4 steps.

@@

 A lot of research has been done on creating ontologies, ranging from what an ontology structure should look like to how ontologies should be generated unsupervised. Some research has focused on ontology merging, with promising but slowing results.

 Recently, a model that embeds words in a vector space that appears to contains semantics has become popular. Since it is unsupervised and can be used to calculate a semantic distance between words, it may very well be applicable to ontology matching. The model will be explained much more in-depth in \ref{w2v}.

@@

\newpage
\section{Context}
% Give an overview of the section
 In this section I will discuss all concepts that are required to understand the study. Firstly, the Semantic Web, the key technology that depends very strongly on ontology matching, will be discussed. Then the concept of the Ontology itself will be addressed. Lastly, current alignment strategies, their problems and other research and technologies will be discussed.
 \subsection{The Semantic Web}
 The Semantic Web, or Web 3.0, is a concept of a Web that enhances the User Experience in ways that the current Web can not provide. Web 2.0, the current Web, models the internet as \textit{hubs} (or sites) with a specific purpose and barely any interconnectivity. For example, Facebook is a hub for sharing content with people you consider friends, but if you want to share that content with more people you will have to go to another hub, like imgur or Pinterest.
The Semantic Web, on the other hand, sees web pages as sources of \textit{information} that can be combined whenever the user needs it. This allows not only humans to browse the internet, but also Artificial Intelligences, which will use this information to aid their users.

@@

 \subsection{Other related research}
 \subsubsection{Word sense representation} \label{w2v}
 Skip-gram models have been widely used to represent words as vectors. This is useful as it is easy to calculate the distance between two vectors, which is a useful property for ontology matching. This vector distance has been shown to be related to the semantic similarity between words, which is very useful for ontology alignment, as parts of one ontology have to be aligned with parts of another ontology depending on their semantic relation.

@@

 Word2vec has been expanded in a number of different ways. For example, word order can be preserved, leading to a similarity measure that is closer to syntax, as syntax defines which words go where in a sentence. This preservation was done by having multiple output layers, one for every word in the context. The research shows that the input does not have to be a bag-of-words representation, which will be useful later on. \cite{ling2015}
 Another interesting development is the multi-sense word2vec, which allows for multiple vector representations per word, depending on the number of different definitions a word has. For example, the word bank represents both the monetary institute and the riverside, which would both have a different representation in this extension. It can differentiate between two different meanings based on the context. This will help with resolving ambiguity, a very important factor in ontology alignment.\cite{multisense}
 \subsection{Used technology}
 \subsubsection{AgreementMaker} \label{agreementmaker}
 AgreementMaker is an ontology matching system which obtained the highest F-measure in 6 of the 7 ontology matching tracks of OAEI 2015. \cite{OAEI2015} The system has also been updated the most recent at the time of writing\cite{agreementmaker}.

 The system uses an extensible framework which allows its users to add new modules to the system. This is very useful for research as it allows the researcher to compare different modules by swapping in just those modules in the framework. The module-swapping technique will also be used in this work, but more of that will be discussed in \ref{Evaluation methods}.

 \subsubsection{Factorie} \label{factorie}

 \newpage
 \section{Problem}

 \subsection{Problem}
 We have seen that when information from multiple domains needs to be integrated so that it is possible to reason over them together, the ontologies of those domains need to be aligned or merged. % Waar?
  Those ontologies contain rules about concepts and concept classes in the shape of triplets, with two concepts and a relation. For example, the rule {\tt Professors are humans} contains the concepts \emph{Professor} and \emph{human} and relationship \emph{being}.

@@

 In summary, ontology aligning is required for multi-domain communication but poses a number of problems that make it difficult for concepts, the building blocks of ontologies, to be matched.
 \subsection{Research questions}
 Can distributed representations of concepts be used to find relationships between concepts in ontologies better than existing terminological matchers?\\\\

 Can distributed representations use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers?\\\\

@@

 \subsection{Hypotheses}
 Since multiple implementations of the algorithm are considered, every hypothesis that follows is considered for every implementation.\\\\

 \begin{center}
 \textit{The distributed representations of concepts matcher can find relationships between concepts in ontologies better than existing terminological matchers.}
 \end{center}
 \begin{center} 
 \textit{The distributed representations matcher can use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers.}
 \end{center}
 \begin{center}
 \textit{The distributed representations matcher can perform well in a domain it was not designed for.}
 \end{center}

@@

 \begin{center}  
 \textit{A distributed representations of concepts matcher can find relationships between concepts in ontologies better than existing terminological matchers.}
 \end{center} 
 \begin{center}
 \textit{A distributed representations matcher can use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers.}
 \end{center}
 \begin{center}
 \textit{A distributed representations matcher can perform well in a domain it was not designed for.}
 \end{center}

@@

 However, ontology matching is also used outside of the semantic web. It is already in use for companies that want to merge knowledge from other companies, for example. If the proposed matching algorithm is better than the current algorithms, or at least improves the performance when used in ensemble with the current algorithms, it will be useful in any situation, not just on the Semantic Web.

 \newpage
 \section{Method}
 In this section, I will describe the method that I developed. Firstly, I will describe the general idea. In the second section, I will give a theoretical justification for why this method should work and how it works conceptually. After that I will describe the different ways that the input graphs can be used to train the word2vec model and create the alignment, some of which I have discarded on theoretical grounds as I will explain in the relevant section. Lastly, I will talk about the implementation process.
 \subsection{Idea}
 As explained earlier, word2vec is a model that learns two things: implicitly it learns the semantic vector representation of words, and explicitly it learns to predict words from their context or vice versa. I will use both of these pieces of information in our alignment algorithm.
 The semantic representation will be used for finding nodes in the two ontology graphs that are similar, i.e. close in the vector space. This step is relatively straightforward if proper vector representations are found. However, this may be hard due to a number of problems which I have listed below.

@@

 \subsection{Discarded implementations}

@@

 \subsection{Implementation}
 The two foundations the algorithm has been build and tested on are Factorie, described in \ref{factorie}, and AgreementMaker, described in \ref{agreementmaker}, respectively. Earlier work on a forked branch of factorie was done by \cite{multisensecode} to implement multi-sense word2vec.

 \newpage
 \section{Evaluation}
 %https://hal.inria.fr/hal-00917910/document
 The goal of the research is to find an ontology alignment algorithm $A$ and proof that it performs better than current techniques. However, as current techniques may combine many different algorithms and sources of information, maybe such a comparison is not fair, nor is a comparison with individual algorithms as they may be optimised for being combined. So, instead we may want to see if performance is improved when adding $A$ to an ensemble of algorithms $E=(A_1,..,A_n)$.
 \subsection{Evaluation methods} \label{Evaluation methods}
 The evaluation should be done using a performance measure $P(Al,T)$ that takes the alignment an algorithm produces given an ontology and the ground truth. If both the alignment and truth are considered sets of alignments, i.e. $Al=\{a_1,..,a_n\}$ and $T=\{a'_1,..,a'_m\}$, then the True Positives $TP=\{a|a\in Al, a\in T\}$, False Positives $FP=\{a|a\in Al, a\notin T\}$ and False Negatives $FN=\{a|a\in T, a\notin Al\}$ determine the performance of the alignment. The True Negatives do not matter, as there are extremely many, and the goal of ontology matching is to find True Positives. Therefore the Jaccard similarity should be used as a performance measure.

@@

 \newpage
 \section{Algorithm}
 \subsection{Idea}
 The goal of any ontology matcher is to find an alignment between two ontologies that is as similar as possible to the true alignment. Such an alignment consists of node pairs where the two nodes in each pair come from the two different ontologies. Sometimes a label is added to each pair, indicating the relationship the two nodes have. For example, 'part-of', 'similar-to', and such. However, we only consider equality relationships and thus the label can be omitted.

@@

 Word2vec is an unsupervised method that maps words to a vector space that has been shown to contain semantic properties. %cite
 It can be trained on large corpora quickly and the resulting vectors can easily be used to measure the distance between words. All these properties make it an excellent candidate for a concept distance measure to use for ontology matching. As an added benefit, it can also take into account the context of a concept, which may be able to improve its performance over a matcher that only uses the label of a concept.
 \subsection{Theory}

@@

 \newpage
 \section{Results} \label{results}

 \newpage
 \section{Conclusions}
 \subsection{Recap research questions}
 \subsection{Recap hypotheses}
 \subsection{Conclusion per hypothesis}
 \subsection{Overview of results}
 \subsection{Final conclusion}

 \newpage
 \section{Discussion}
 \subsection{Interpretation}
 \subsection{Unexpected results}
 \subsection{Expected results}
 \subsection{Future research}

 \newpage
 \bibliographystyle{plain}
 \bibliography{bibliography}{}

 \newpage
 \begin{appendices}
  \pagenumbering{Roman}
  \section{Appendix 1}
  \section{Appendix 2}
 \end{appendices}

 \end{document}
